The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-09 rational.tex
algebra with wire superimposed

To get a feeling for the main idea, consider the following algebraic
manipulation relating a natural number $a$ to itself:
\[\begin{array}{rclll}
&& a & \mbox{initial~$a$} & (0) \\
&=& a + 0 & \mbox{$0$~is~identity~for~$+$} & (1) \\
&=& a + (-a + a) & \mbox{$-a$~is~the~additive~inverse~of~$a$} & (2) \\
&=& (a + (-a)) + a & \mbox{$+$~is~associative} & (3) \\
&=& 0 + a & \mbox{$-a$~is~the~additive~inverse~of~$a$} & (4) \\
&=& a & \mbox{$0$~is~identity~for~$+$} & (5)
\end{array}\]
\[\begin{array}{rlllllllll}
 \!&\!\!  a \!&\! \!&\!       \!&\! \!&\!    \!&\! \mbox{initial~$a$} \!&\! (0) \\
=\!&\!\!  a \!&\!+\!&\!       \!&\!0\!&\!    \!&\! \mbox{$0$~is~identity~for~$+$} \!&\! (1) \\
=\!&\!\!  a \!&\!+\!&\! (-a   \!&\!+\!&\! a) \!&\! \mbox{$-a$~is~the~additive~inverse~of~$a$} \!&\! (2) \\
=\!&\!\! (a \!&\!+\!&\! (-a)) \!&\!+\!&\! a  \!&\! \mbox{$+$~is~associative} \!&\! (3) \\
=\!&\!\!    \!&\!0\!&\!       \!&\!+\!&\! a  \!&\! \mbox{$-a$~is~the~additive~inverse~of~$a$} \!&\! (4) \\
=\!&\!\!    \!&\! \!&\!       \!&\! \!&\! a  \!&\! \mbox{$0$~is~identity~for~$+$} \!&\! (5)
\end{array}\]

>>>>

To get a feeling for the main idea, consider the following algebraic
manipulation relating a natural number $a$ to itself:
%% \[\begin{array}{rclll}
%% && a & \mbox{initial~$a$} & (0) \\
%% &=& a + 0 & \mbox{$0$~is~identity~for~$+$} & (1) \\
%% &=& a + (-a + a) & \mbox{$-a$~is~the~additive~inverse~of~$a$} & (2) \\
%% &=& (a + (-a)) + a & \mbox{$+$~is~associative} & (3) \\
%% &=& 0 + a & \mbox{$-a$~is~the~additive~inverse~of~$a$} & (4) \\
%% &=& a & \mbox{$0$~is~identity~for~$+$} & (5)
%% \end{array}\]
%% \[\begin{array}{rlllllllll}
%%  \!&\!\!  a \!&\! \!&\!       \!&\! \!&\!    \!&\! \mbox{initial~$a$} \!&\! (0) \\
%% =\!&\!\!  a \!&\!+\!&\!       \!&\!0\!&\!    \!&\! \mbox{$0$~is~identity~for~$+$} \!&\! (1) \\
%% =\!&\!\!  a \!&\!+\!&\! (-a   \!&\!+\!&\! a) \!&\! \mbox{$-a$~is~the~additive~inverse~of~$a$} \!&\! (2) \\
%% =\!&\!\! (a \!&\!+\!&\! (-a)) \!&\!+\!&\! a  \!&\! \mbox{$+$~is~associative} \!&\! (3) \\
%% =\!&\!\!    \!&\!0\!&\!       \!&\!+\!&\! a  \!&\! \mbox{$-a$~is~the~additive~inverse~of~$a$} \!&\! (4) \\
%% =\!&\!\!    \!&\! \!&\!       \!&\! \!&\! a  \!&\! \mbox{$0$~is~identity~for~$+$} \!&\! (5)
%% \end{array}\]

\begin{center}
\scalebox{1.1}{
\includegraphics{diagrams/thesis/algebra-wire1.pdf}
}
\end{center}
